---
title: ''
author: ''
date: ''
output: 
  html_document:
    df_print: paged
    keep_md: true
  word_document:
    reference_docx: style.1.docx
---

```{r setup, include = FALSE}

  knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                        fig.width  = 6 * 1.67, fig.height = 6)

  load("data/data.RData")
  
  library(descr)
  library(tidyverse)
  library(viridis)
  library(flextable)
  library(patchwork)
  library(rms)
  library(Hmisc)
  library(knitr)
  library(scales)
  library(viridis)
  library(RColorBrewer)

```


```{r var_plot}

  var_plot <- function(data, var, log = FALSE){

    df <- data

    if(log == TRUE){

      df <- filter(df, type == "Healthy") %>%
        select(id, {{var}}) %>%
        group_by(id) %>%
        summarise(mean_1 = mean(log({{var}}), na.rm = TRUE)) %>%
        full_join(df, by = "id")

      df <- filter(df, type == "Cancer") %>%
        select(id, {{var}}) %>%
        group_by(id) %>%
        summarise(mean_2 = mean(log({{var}}), na.rm = TRUE)) %>%
        full_join(df, by = "id")

      g1 <- mutate(df, id = reorder(id, mean_1)) %>%
        ggplot(aes(color = log(frequency), y = log({{var}}), x = id,
                   group = id)) +
        geom_point(size = 1, alpha = 0.5) +
        scale_color_viridis("log(Freq)", trans = "reverse") +
        geom_point(aes(y = mean_1), shape = "square", color = "white") +
        geom_point(aes(y = mean_2), shape = "square", color = "red") +
        facet_wrap(~type, nrow = 1) +
        theme(panel.background = element_rect(fill = "grey40"),
              panel.grid = element_blank(), 
              axis.text.x = element_text(angle = 90, hjust = 1, size = 6)) +
        xlab("")

    } else{

      df <- filter(df, type == "Healthy") %>%
        select(id, {{var}}) %>%
        group_by(id) %>%
        summarise(mean_1 = mean({{var}}, na.rm = TRUE)) %>%
        full_join(df, by = "id")

      df <- filter(df, type == "Cancer") %>%
        select(id, {{var}}) %>%
        group_by(id) %>%
        summarise(mean_2 = mean({{var}}, na.rm = TRUE)) %>%
        full_join(df, by = "id")

      g1 <- mutate(df, id = reorder(id, mean_1)) %>%
        ggplot(aes(color = log(frequency), y = {{var}}, x = id, group = id)) +
        geom_point(size = 1, alpha = 0.5) +
        scale_color_viridis(guide = FALSE, "log(Freq)", trans = "reverse") +
        geom_point(aes(y = mean_1), shape = "square", color = "white") +
        geom_point(aes(y = mean_2), shape = "square", color = "red") +
        facet_wrap(~type, nrow = 1) +
        theme(panel.background = element_rect(fill = "grey40"),
              panel.grid = element_blank(), 
              axis.text.x = element_text(angle = 90, hjust = 1, size = 6)) +
        xlab("")
    }
    return(g1)
  }


```

# Data

## Pre-processing

Patient 15 is the only patient with 2 sets of cancerous tissue measures, and
both of them are much much higher in terms of impedance values at the lowest
frequencies (demonstrated below, in comparison to two other randomly selected patients; note the log10 scale - these are very large differences). I have thus removed patient 15 for now (though the fact that both cancer samples appear unusual relative to the other patients may bear some consideration later). 

```{r}

  data[grepl("15", data$id) | data$id %in% c(1, 6), ] %>%
  ggplot(aes(y = imaginary_z, x = frequency, group = patient_id, 
             color = type)) +
  geom_line(size = 1, alpha = 0.5) +
  facet_wrap(~id) +
  scale_x_continuous(labels = function(x) format(x, scientific = TRUE)) +
  scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
    ) +
  theme(axis.text.x = element_text(size = 6)) +
  scale_color_brewer(guide = FALSE, palette = "Set1") 

```

```{r}

  data <- filter(data, id != 15)

```

## Description 

After the removal of this patient, the dataset now contains information on `r length(unique(data$id))` patients, each with a set of measurements taken at 50 different frequencies on both cancerous and healthy tissue (`r length(unique(data$id))` patients X `r length(unique(data$frequency))` frequencies X `r length(levels(data$type))` tissue types = 2000 total measurements). 


### Key output values by patient and frequency

#### Imaginary Z

Values for each patient

```{r}
  
  ggplot(
    data, 
    aes(x = frequency, y = imaginary_z, group = interaction(id, type), 
        color = type)
    ) + 
  geom_line(size = 1, alpha = 0.5) +
  facet_wrap(~id) +
  scale_x_continuous(labels = function(x) format(x, scientific = TRUE)) +
  scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
    ) +
  theme(axis.text.x = element_text(size = 6)) +
  scale_color_brewer("", palette = "Set1") 

```

Superimposed values for each patient

```{r}
  
  ggplot(
    data, 
    aes(x = frequency, y = imaginary_z, group = interaction(id, type), 
        color = id)
    ) + 
  geom_line(size = 1, alpha = 0.5) +
  facet_wrap(~type) +
  scale_x_continuous(labels = function(x) format(x, scientific = TRUE)) +
  scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
    ) +
  theme(axis.text.x = element_text(size = 6)) +
  scale_color_manual(guide = FALSE, values = rep(brewer.pal(5, "Set1"), 4))

```

Cancer value minus healthy tissue value (note: y axis truncated at +/- 10^05)

```{r}

  distinct(data, id, frequency, imaginary_z_diff) %>%
  ggplot(
    aes(x = frequency, y = imaginary_z_diff, group = id, color = id)
    ) + 
  geom_line(size = 1, alpha = 0.5) +
  geom_hline(yintercept = 0) +
  coord_cartesian(ylim = c(-100000, 100000)) +
  scale_x_continuous(labels = function(x) format(x, scientific = TRUE)) +
  theme(axis.text.x = element_text(size = 6)) +
  scale_color_manual(guide = FALSE, values = rep(brewer.pal(5, "Set1"), 4))

```

Distribution of values by frequency and tissue type

```{r}

  ggplot(data, aes(y = imaginary_z, x = interaction(type, frequency), 
                   fill = type, color = type)) +
    geom_violin() +
    scale_fill_brewer("", palette = "Set1") +
    scale_color_brewer(guide = FALSE, palette = "Set1") +
    xlab("") +
    scale_y_log10() +
    theme(axis.text.x = element_blank(), 
          panel.grid = element_blank()) 

```

```{r}

  g1 <- var_plot(data, imaginary_z)
  g2 <- var_plot(data, imaginary_z, TRUE)
  
  g1 + g2

```


#### Real Z

Values for each patient 

```{r}

  ggplot(
    data, 
    aes(x = frequency, y = real_z, group = interaction(id, type), 
        color = type)
    ) + 
  geom_line(size = 1, alpha = 0.5) +
  facet_wrap(~id) +
  scale_x_continuous(labels = function(x) format(x, scientific = TRUE)) +
  scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
    ) +
  theme(axis.text.x = element_text(size = 6)) +
  scale_color_brewer("", palette = "Set1") 

```

Superimposed values for each patient

```{r}

  ggplot(
    data, 
    aes(x = frequency, y = real_z, group = interaction(id, type), 
        color = id)
    ) + 
  geom_line(size = 1, alpha = 0.5) +
  facet_wrap(~type) +
  scale_x_continuous(labels = function(x) format(x, scientific = TRUE)) +
  scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
    ) +
  theme(axis.text.x = element_text(size = 6)) +
  scale_color_manual(guide = FALSE, values = rep(brewer.pal(5, "Set1"), 4))

```

Cancer value minus healthy tissue value (note: y axis truncated at +/- 10^05)

```{r}

  distinct(data, id, frequency, real_z_diff) %>%
  ggplot(
    aes(x = frequency, y = real_z_diff, group = id, color = id)
    ) + 
  geom_line(size = 1, alpha = 0.5) +
  geom_hline(yintercept = 0) +
  coord_cartesian(ylim = c(-100000, 100000)) +
  scale_x_continuous(labels = function(x) format(x, scientific = TRUE)) +
  theme(axis.text.x = element_text(size = 6)) +
  scale_color_manual(guide = FALSE, values = rep(brewer.pal(5, "Set1"), 4))

```

Distribution of values by frequency and tissue type

```{r}

  ggplot(data, aes(y = real_z, x = interaction(type, frequency), 
                   fill = type, color = type)) +
    geom_violin() +
    scale_fill_brewer("", palette = "Set1") +
    scale_color_brewer(guide = FALSE, palette = "Set1") +
    xlab("") +
    scale_y_log10() +
    theme(axis.text.x = element_blank(), 
          panel.grid = element_blank()) 

```

```{r}

  g1 <- var_plot(data, real_z)
  g2 <- var_plot(data, real_z, TRUE)
  
  g1 + g2

```


#### Magnitude z

Values for each patient

```{r}

  ggplot(
    data, 
    aes(x = frequency, y = magnitude_z, group = interaction(id, type), 
        color = type)
    ) + 
  geom_line(size = 1, alpha = 0.5) +
  facet_wrap(~id) +
  scale_x_continuous(labels = function(x) format(x, scientific = TRUE)) +
  scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
    ) +
  theme(axis.text.x = element_text(size = 6)) +
  scale_color_brewer("", palette = "Set1") 

```

Superimposed values for each patient 

```{r}

  ggplot(
    data, 
    aes(x = frequency, y = magnitude_z, group = interaction(id, type), 
        color = id)
    ) + 
  geom_line(size = 1, alpha = 0.5) +
  facet_wrap(~type) +
  scale_x_continuous(labels = function(x) format(x, scientific = TRUE)) +
  scale_y_log10(
    breaks = scales::trans_breaks("log10", function(x) 10^x),
    labels = scales::trans_format("log10", scales::math_format(10^.x))
    ) +
  theme(axis.text.x = element_text(size = 6)) +
  scale_color_manual(guide = FALSE, values = rep(brewer.pal(5, "Set1"), 4))

```

Cancer value minus healthy tissue value (note: y axis truncated at +/- 10^05)

```{r}

  distinct(data, id, frequency, magnitude_z_diff) %>%
  ggplot(
    aes(x = frequency, y = magnitude_z_diff, group = id, color = id)
    ) + 
  geom_line(size = 1, alpha = 0.5) +
  geom_hline(yintercept = 0) +
  coord_cartesian(ylim = c(-100000, 100000)) +
  scale_x_continuous(labels = function(x) format(x, scientific = TRUE)) +
  theme(axis.text.x = element_text(size = 6)) +
  scale_color_manual(guide = FALSE, values = rep(brewer.pal(5, "Set1"), 4))

```

Distribution of values by frequency and tissue type

```{r}

  ggplot(data, aes(y = magnitude_z, x = interaction(type, frequency), 
                   fill = type, color = type)) +
    geom_violin() +
    scale_fill_brewer("", palette = "Set1") +
    scale_color_brewer(guide = FALSE, palette = "Set1") +
    xlab("") +
    scale_y_log10() +
    theme(axis.text.x = element_blank(), 
          panel.grid = element_blank()) 

```


# Prediction models

For now let's just so a simple logit and ignore the repeat samples. 

```{r}

  m1 <- glm(type ~ rcs(real_z, 5) * rcs(imaginary_z, 5), 
            data = data, family = binomial)
 
  data$m1_pred <- predict(m1, type = "response")
  
  val_m1 <- val.prob(data$m1_pred, as.numeric(data$type) - 1, 
                     pl = FALSE) %>% round(3)
  
  rescale_brier <- function(x, p, ...){ 
    format(round(1 - (x / (mean(p) * (1 - mean(p)))), digits = 2), nsmall = 2)
  }
  
  b1 <- rescale_brier(val_m1["Brier"], 0.512) 

```


```{r}

  kable(val_m1)

```

```{r}

  ggplot(data, aes(x = log(real_z), y = log(imaginary_z), color = m1_pred)) +
    geom_point() +
    scale_color_viridis()

```





```{r code_book}

  print(summarytools::dfSummary(data), method = "render")

```

```{r sysinfo}

  DescTools::SysInfo()

```
